A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2022-12-28 paper/fascd.tex
edits section 8

\section{Full approximation scheme constraint decomposition (\fascd) V-cycle} \label{sec:vcycle}

The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  \fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}.  Note that we coarsen both the residual and the solution iterate itself, in constrast to classical linear multigrid wherein only the residual is coarsened.  In fact, \fascd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).

The V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$, but we will use prolongation $P$ to indicate where the computer representation changes.

However, the corrected iterate needs to be admissible.  In detail, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing, an incomplete solution process on a coarser-level VI problem (defined below), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  At the bottom the coarsest-level VI is then solved, resulting in a downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.  Smoothing in $\mathcal{U}^1$ yields admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.  Thus the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.

@@

To justify \fascd, namely Algorithm \ref{alg:nmcd} below, we must state the problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, along with creating a source functional.  Recall that Table \ref{tab:transfers} summarizes the needed transfer operators.

Let $j$ denote the mesh level of the coarsened problem.  We assume a (re-)discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, using quadrature and etc., presumably by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.

The formulas are as follows.  Finest-level problem \eqref{eq:fe:vi} is defined by the discretized operator $f^J$ and the original source functional $\ell^J$.  A finest-level, admissible iterate $w^J \approx u^J$, $w^J \in \mathcal{K}^J$, has already\footnote{Conceptually-speaking, at least.  In practice one generates LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ during the descent stage; see Algorithm \ref{alg:nmcd}.  Observe that the LDCs depend neither on coarsened iterates $g^j$ nor corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j,\mathcal{U}^j$ at each level (Section \ref{sec:femultilevel}).  As we descend in the V-cycle, we recursively define $j$th-level iterates
\begin{equation}
g^j = \begin{cases} w^J, & j=J \\
                    \iR(g^{j+1} + y^{j+1}), & j < J.
      \end{cases}  \label{eq:fe:defineg}
\end{equation}
These are the iterates \emph{before} smoothing and/or correction has been applied at level $j$.  (The problem solved by $y^{j+1}$ is stated momentarily.)  From $g^j$ we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(g^j) + R\left(\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
Observe that injection $\iR$ coarsens the iterates $g^j$---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---while canonical restriction $R$ coarsens the source functionals.

On descent, $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(g^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
\end{equation}
for which the initial iterate is
\begin{equation}
(y^j)^{(0)}=0.  \label{eq:fe:downwardinitial}
\end{equation}
(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved, but in a larger admissible set.  We find $z^j \in \mathcal{U}^j$ so that
\begin{equation}
\ip{f^j(g^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
\end{equation}
The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the current (down-smoothed) iterate on the level:\footnote{Prolongation $P$ is used in \eqref{eq:fe:upwardinitial} to reflect the change in mesh/FE representation, from $\mathcal{T}^{j-1}$ to $\mathcal{T}^{j}$.}
\begin{equation}
(z^j)^{(0)} = y^j + P z^{j-1}.  \label{eq:fe:upwardinitial}
\end{equation}
For the $j=0$ coarsest level, $(z^0)^{(0)}=0$.  Initial iterate \eqref{eq:fe:upwardinitial} is admissible since $\mathcal{U}^j = \mathcal{D}^j + \mathcal{U}^{j-1}$; see \eqref{eq:fe:constraintdecomp} or \eqref{eq:fe:upwardsumcd}.

If we were to inject the original bi-obstacles from \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:defineg} that $\underline{\gamma}^j \le g^j \le \overline{\gamma}^j$, just like $\underline{\gamma}^J \le w^J \le \overline{\gamma}^J$ for the finest-level iterate.  In this narrow sense $g^j$ is admissible with respect to the original (finest-level) box constraints.  (In fact, while $g^j \in \cV^j \subset \cV^J$ certainly holds, generally $g^j \notin \cK^J$.)  However, following \cite{GraeserKornhuber2009}, our use of defect constraints (Table \ref{tab:ldcs}) avoids the need to generate functions $\underline{\gamma}^j$ and $\overline{\gamma}^j$, and it yields a final, finest-level iterate $g^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.

>>>>

\section{Full approximation storage constraint decomposition (\fascd) V-cycle} \label{sec:vcycle}

The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  \fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).

The V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$, but we will use prolongation $P$ to indicate where the computer representation changes.  Note that we coarsen both the residual and the solution iterate, in constrast to classical linear multigrid wherein only the residual is coarsened.

However, the corrected iterate needs to be admissible.  In detail, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing, an incomplete solution process on a VI problem (Section \ref{sec:smoothers}), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  At the bottom the coarsest-level VI is then solved (also Section \ref{sec:smoothers}), resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.  Smoothing in $\mathcal{U}^1$ yields admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.  Then the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.

@@

To state \fascd Algorithm \ref{alg:nmcd} below we must clarify the problems solved at each mesh level.  As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.  On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, presumably using quadrature and etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.  Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$, and an admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{Conceptually-speaking this is true.  In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:nmcd}.  Observe that the LDCs do not depend on $w^j$ for $j<J$, nor on the corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:ldcs}).

As we descend $j<J$, we define new $j$th-level iterates
\begin{equation}
w^j = \iR(w^{j+1} + y^{j+1}),  \label{eq:fe:definew}
\end{equation}
the value \emph{before} smoothing and/or correction at level $j$.  From $w^j$ we define $\ell^j \in (\cV^j)'$:
\begin{equation}
\ell^j = \begin{cases} \ell^J, & j=J \\
                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
\end{equation}
Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---while canonical restriction $R$ coarsens linear functionals.  (Recall Table \ref{tab:transfers} summarizes our transfer operators.)

On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
\begin{equation}
\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
\end{equation}
for which the initial iterate is $y_{(0)}^j=0$.  (Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved, but in a larger admissible set, for $z^j \in \mathcal{U}^j$:
\begin{equation}
\ip{f^j(w^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
\end{equation}
The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the current (down-smoothed) correction on the level,\footnote{Prolongation $P$ is used in \eqref{eq:fe:upwardinitial} to reflect the change in FE representation, from $\mathcal{V}^{j-1}$ to $\mathcal{V}^{j}$.}
\begin{equation}
z_{(0)}^j = y^j + P z^{j-1},  \label{eq:fe:upwardinitial}
\end{equation}
but with $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible since $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$; see \eqref{eq:fe:upwardsuminclusion}.

If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.  In this sense $w^j$ is admissible with respect to the box constraints, but, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate such box constraints $\underline{\gamma}^j$ and $\overline{\gamma}^j$ on each level.  (Note that while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$.)  In any case, Algorithm \ref{alg:nmcd} will yield a final, finest-level iterate $w^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
